---
title: "MDA-Milestone-1"
output: github_document
---

# Task 1
### Task 1.1
1. cancer_sample
2. parking_meters
3. steam_games
4. vancouver_trees

### Task 1.2
`````{r}
#Exploring datasets of interest. For each dataset, the below code provides class type, number of rows, number of columns, variable names, and provides information on how data is provided (ie. numeric or characters)

#cancer_sample
class(cancer_sample)
glimpse(cancer_sample)

#parking_meters
class(parking_meters)
glimpse(parking_meters)

#steam_games
class(steam_games)
glimpse(steam_games)

#vancouver_trees
class(vancouver_trees)
glimpse(vancouver_trees)
`````

### Task 1.3
1. cancer_sample
2. steam_games

I selected these two datasets because they contain many variables that I believe could be used to figure out some interesting information. They both provide a relatively large number of variables, so many different graphs can be made to visualize potential relations or trends. For example, when looking at the cancer_sample dataset, it would be interesting to see if there are any correlations between different traits of the tumor (ie.mean tumor compactness and mean tumor area). This could be done by creating a scatter plot and adding a trendline. In terms of the steam_games dataset, it could be interesting to see if the genre of a game is related to the sale price, or if theres any relationship between genre and mature content. 

### Task 1.4
The research question I would ask for the cancer_sample dataset would be - Are there any correlations that can be found between the different traits of a tumor, and do these traits relate to the diagnosis? 

The research question I would ask for the steam_games dataset would be - Are there any correlations between the maturity rating and genre?

I've decided to work with the cancer_sample dataset for the remainder of this milestone

# Task 2
### Task 2.1
##### 2.1.1
`````{r}
#selects cancer_sample dataset and produces a graph with area_mean on the x-axis and perimeter_mean on the y-axis.
cancer_sample %>%
  ggplot(aes(area_mean,perimeter_mean)) +
#plots the points for the above two vaiables, along with added a colour to the points to differentiate between the diagnosis of the tumor. An alpha of 0.2 was used to be able to see areas of overlap.
  geom_point(aes(colour=diagnosis), alpha=0.2) +
#sets the origin of the graph to 0,0 and ensures all points are included in the plot.
  expand_limits(x = 0, y = 0) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 2600)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 200))
`````
##### 2.1.6
`````{r}
#selects cancer_sample dataset and produces a graph with diagnosis on the x-axis and compactness_mean on the y-axis.
cancer_sample %>%
  ggplot(aes(diagnosis,compactness_mean)) +
#produces a boxplot and makes the width of the boxplot smaller.
  geom_boxplot(width = 0.5) +
#adds the individuals data points, makes the points' transparency lower and smaller, and shifts the points horizontally while keeping the vertical position the same.
  geom_point(alpha = 0.2, size = 0.2, position = position_dodge2(w = 0.25))
`````
##### 2.1.7
`````{r}
#selects all variables from ID to fractal_dimension_mean and produces a new tibble including only this information.
select(cancer_sample, ID:fractal_dimension_mean)
`````
##### 2.1.8
`````{r}
#selects cancer_sample dataset and produces a graph with area_mean on the x_axis.
cancer_sample %>%
  ggplot(aes(area_mean)) +
#produces a density graph containing two lines that represent the two different types of diagnoses with the area below each line filled. Alpha was used to add transparency to the the colour filled area.
  geom_density(aes(fill = diagnosis), alpha = 0.3) +
#set origin of graph to 0,0
  expand_limits(x = 0, y = 0) +
#relablled y-axis from y, to density
  ylab(label = "density")
`````

### Task 2.2

# Task 3
