---
title: "mda-Milestone-2"
output: html_document
---

`````{r}
#loads needed packages to execute the code in this file and supresses startup messages
suppressPackageStartupMessages(library(datateachr)) #allows for access to the datasets in this file.
suppressPackageStartupMessages(library(tidyverse)) # loads packages needed for data visualization (ggplot2), and data manipulation (dplyr) that were used in this file.
`````

# Task 1
### Task 1.1
1. What correlation exists (if any) between the symmetry of a tumor and texture of a tumor?*
2. Is there a correlation between tumor diagnosis and tumor smoothness?**
3. Is there a correlation between the area and perimeter of a tumor in regards to tumor diagnosis?*
4. Is there a correlation between the tumor's fractal dimension and compactness in regards to tumor diagnosis?**

Changes from last milestone:
*Wording of question was changed to ask about correlation instead or relationship between variables
**Research Question was changed to be less broad

### Task 1.2
####Research Question 1
###### Summarizing
```{r}
(cancer_sample %>%
  group_by(diagnosis) %>%
  summarise(symmetry_geometric_mean = geometric.mean(symmetry_mean, na.rm = TRUE), symmetry_range = paste("(", min(symmetry_mean, na.rm=T), "-", max(symmetry_mean,na.rm=T), ")"), symmetry_median = median(symmetry_mean, na.rm = TRUE), symmetry_count = n()))
```

###### Graphing
```{r}

```

####Research Question 2 (q1)
###### Summarizing
```{r}
(cancer_sample %>%
  group_by(diagnosis) %>%
  summarise(overall_smoothness_mean = mean(smoothness_mean, na.rm = TRUE), smoothness_range = paste("(", min(smoothness_mean, na.rm=T), "-", max(smoothness_mean,na.rm=T), ")"), smoothness_median = median(smoothness_mean, na.rm = TRUE), smoothness_count = n()))
```

###### Graphing
```{r}

  
```

####Research Question 3 (2)
###### Summarizing
```{r}
cancer_sample %>%
  group_by(diagnosis) %>%
  mutate(relative_tumor_size = case_when(area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 TRUE ~ "large")) %>%
  select(ID:fractal_dimension_mean, relative_tumor_size)
```

###### Graphing
```{r}

```

####Research Question 4
###### Summarizing
```{r}
(cancer_sample %>%
  group_by(diagnosis) %>%
  summarise(total_number_of_tumors = n()))
```

###### Graphing
```{r}

```

### Task 1.3


# Task 2
### Task 2.1


### Task 2.2


### Task 2.3


# Additional Comments