---
title: "mda-Milestone-2"
output: html_document
---

`````{r}
#loads needed packages to execute the code in this file and supresses startup messages
suppressPackageStartupMessages(library(datateachr)) #allows for access to the datasets in this file.
suppressPackageStartupMessages(library(tidyverse)) # loads packages needed for data visualization (ggplot2), and data manipulation (dplyr) that were used in this file.
`````

# Task 1
### Task 1.1
1. What correlation exists (if any) between the symmetry of a tumor and texture of a tumor?*
2. Is there a correlation between tumor diagnosis and tumor smoothness?**
3. Is there a correlation between the area and perimeter of a tumor in regards to tumor diagnosis?*
4. Is there a correlation between the tumor's fractal dimension and tumor diagnosis?**

Changes from last milestone:
*Wording of question was changed to ask about correlation instead or relationship between variables
**Research Question was changed to be less broad

### Task 1.2
####Research Question 1
###### Summarizing (option 1)
```{r}
(cancer_sample %>%
  group_by(diagnosis) %>%
  summarise(symmetry_geometric_mean = geometric.mean(symmetry_mean, na.rm = TRUE), symmetry_range = paste("(", min(symmetry_mean, na.rm=T), "-", max(symmetry_mean,na.rm=T), ")"), symmetry_median = median(symmetry_mean, na.rm = TRUE), symmetry_count = n()))
```

###### Graphing (option 7)
```{r}
#selects cancer_sample dataset and produces a graph with symmetry_mean on the x-axis and texture_mean on the y-axis.
cancer_sample %>%
  ggplot(aes(symmetry_mean,texture_mean)) +
  
#plots the points for the above two variables with an alpha of 0.2 to allow for visualization of overlapping data points. A colour aesthetic was also used to show the diagnosis of each point.
  geom_point(alpha = 0.2) +
 
#sets the origin of the graph to 0,0 and ensures all points are included in the plot.
  expand_limits(x = 0, y = 0) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, max(cancer_sample$symmetry_mean)+0.05)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(cancer_sample$texture_mean)+5)) +
  
  #adds a linear trendline
  stat_smooth(method = 'lm', aes(colour = "red"), se = FALSE, fullrange=TRUE) +
  
  #removes legend produced from trendline 
  theme(legend.position="none")
```

####Research Question 2 
###### Summarizing (option 1)
```{r}
(cancer_sample %>%
  group_by(diagnosis) %>%
  summarise(overall_smoothness_mean = mean(smoothness_mean, na.rm = TRUE), smoothness_range = paste("(", min(smoothness_mean, na.rm=T), "-", max(smoothness_mean,na.rm=T), ")"), smoothness_median = median(smoothness_mean, na.rm = TRUE), smoothness_count = n()))
```

###### Graphing (option 7) 
```{r}
#selects cancer_sample dataset and produces a graph with diagnosis on the x-axis and smoothness_mean on the y-axis.
cancer_sample %>%
  ggplot(aes(diagnosis,smoothness_mean)) +
  
#produces a boxplot and makes the width of the boxplot smaller.
  geom_boxplot(width = 0.5) +
  
#adds the individuals data points, makes the points' transparency lower and smaller, and shifts the points horizontally while keeping the vertical position the same.
  geom_jitter(width = 0.2, size = 0.3, alpha = 0.2) +
  
#makes the y-axis start ar origin, and makes sure all points are within the graph  
  expand_limits(x = 0, y = 0) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(cancer_sample$smoothness_mean)+0.03)) 
```

####Research Question 3
###### Summarizing (option 3) 
```{r}
cancer_sample %>%
  group_by(diagnosis) %>%
  mutate(relative_tumor_size = case_when(area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 TRUE ~ "large")) %>%
  select(ID:fractal_dimension_mean, relative_tumor_size)
```

###### Graphing (option 6)
```{r}
#selects cancer_sample dataset and produces a graph with area_mean on the x-axis and perimeter_mean on the y-axis.
cancer_sample %>%
  ggplot(aes(area_mean,perimeter_mean)) +
  
#plots the points for the above two variables with an alpha of 0.2 to allow for visualization of overlapping data points. A colour aesthetic was also used to show the diagnosis of each tumor.
  geom_point(aes(colour = diagnosis), alpha = 0.2) +

#makes axes logarithmic
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10") +
  
#adds log ticks for better analysis
  annotation_logticks() +
  
#adds logarithmic trend line
  stat_smooth(method="lm",formula=y~log(x), colour="black")
```

####Research Question 4
###### Summarizing (option 2)
```{r}
(cancer_sample %>%
  group_by(diagnosis) %>%
  summarise(total_number_of_tumors = n()))
```

###### Graphing (option 7) density
```{r}
#selects cancer_sample dataset and produces a graph with fractal_dimension_mean on the x_axis.
cancer_sample %>%
  ggplot(aes(fractal_dimension_mean)) +
  
#produces a density graph containing two lines that represent the two different types of diagnoses with the area below each line filled. Alpha was used to add transparency to the the colour filled area.
  geom_density(aes(fill = diagnosis), alpha = 0.3) +
  
#set origin of graph to 0,0
  expand_limits(x = 0, y = 0) +
  
#relablled y-axis from y, to density.
  ylab(label = "density")
```

### Task 1.3


# Task 2
### Task 2.1


### Task 2.2


### Task 2.3


# Additional Comments