---
title: "mda-Milestone-3"
output: github_document
---

`````{r}
#loads needed packages to execute the code in this file and supresses startup messages
suppressPackageStartupMessages(library(datateachr)) #allows for access to the datasets in this file.
suppressPackageStartupMessages(library(tidyverse)) #loads packages needed for data visualization (ggplot2), and data manipulation (dplyr) that were used in this file.
suppressPackageStartupMessages(library(broom))
```

# Research Questions:

1. Is there a correlation between the area and perimeter of a tumor in regards to tumor diagnosis?
2. Is there a correlation between the tumor's fractal dimension and tumor diagnosis?

# Exercise 1

Had to make a new graph since none of my graphs in Milestone 2 had 3 factors. Below is the code to produce this original graph.

```{r}
#selects cancer_sample dataset
cancer_sample %>%
  
#produces an additional variable with 5 possible categories as the observations based on area_mean  
  mutate(relative_tumor_size = case_when(area_mean < 200 ~ "very low",                                    
                                 area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ " very large")) %>%
  
#produces a graph with relative_tumor_size on the x-axis and area_mean on the y-axis.  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
#produces a boxplot and makes the width of the boxplot smaller.  
  geom_boxplot(width = 0.5) +
  
#adds the individuals data points, makes the points' transparency lower and smaller, and shifts the points horizontally while keeping the vertical position the same. Also colours the point based on tumor diagnosis.
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3)
```

#### Task Number: 1

```{r}
#selects cancer_sample dataset
cancer_sample %>%
  
#produces an additional variable with 5 possible categories as the observations based on area_mean  
  mutate(relative_tumor_size = factor(case_when(area_mean < 200 ~ "very small",                                                                                        area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ "very large"),
                                 
#reorders the factors
                                 levels = c("very small", "small", "medium",  "large", "very large"))) %>%
  
#produces a graph with relative_tumor_size on the x-axis and area_mean on the y-axis.    
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
#produces a boxplot and makes the width of the boxplot smaller.  
  geom_boxplot(width = 0.5) +
  
#adds the individuals data points, makes the points' transparency lower and smaller, and shifts the points horizontally while keeping the vertical position the same. Also colours the point based on tumor diagnosis.  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3)
```

I chose to do this ordering of the factors because I believe when looking at difference in sizes, it seems to be more easy on the eye when you go from the smallest value to the largest. It is also easier to view and see trends because its usually how this kind of measurements are ordered. For example, if I were to remove the axis labels and factor names, I would automatically think that in the original graph, the box plots went from smallest to largest, which it did not. 

#### Task Number: 2

```{r}
#selects cancer_sample dataset
cancer_sample %>%
  
#produces an additional variable with 5 possible categories as the observations based on area_mean  
  mutate(relative_tumor_size = factor(case_when(area_mean < 200 ~ "very small",                                                                                        area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ "very large"),
                                 levels = c("very small", "small", "medium",  "large", "very large"))) %>%
  
#groups 'very small' and 'small' into one factor called 'other'
  mutate(relative_tumor_size = fct_collapse(relative_tumor_size, 
              medium = "medium", large = "large", very_large = "very large", other = c("very small", "small"))) %>%

#produces a graph with relative_tumor_size on the x-axis and area_mean on the y-axis.  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
#produces a boxplot and makes the width of the boxplot smaller.    
  geom_boxplot(width = 0.5) + 
  
#adds the individuals data points, makes the points' transparency lower and smaller, and shifts the points horizontally while keeping the vertical position the same. Also colours the point based on tumor diagnosis.    
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3) +
  
#renames the factor labels so that the 'verylarge' factor name is changed to 'very large'  
  scale_x_discrete(labels = c("other", "medium", "large", "very large"))
```

I chose to do this grouping because I was mostly interested in the medium to largest factors, but did want to remove any available data. By doing this group it also emphasizes how there are very little numbers of benign tumors, and they dont go past the medium factor. This provides information on tumor diagnosis and area_mean. 

# Exercise 2

### Exercise 2.0

Research Question: Is there a correlation between the area and perimeter of a tumor in regards to tumor diagnosis?
Variable of interest: area_mean

### Exercise 2.1
```{r}
#creates a linear model
trendline <- lm(perimeter_mean ~ area_mean, cancer_sample)

#provides information on the output lm() creates
unclassed <- unclass(trendline)
print(unclassed)
```

I used the lm() function to create a linear model, then used the unclass() funciton to look at all the information lm() created. The below code chunk produces a scatterplot between area_mean and perimeter_mean, with a linear regression line.

```{r}
#selects cancer_sample dataset and produces a graph with area_mean on the x-axis and perimeter_mean on the y-axis.
cancer_sample %>%
  ggplot(aes(x = area_mean, y = perimeter_mean)) +
  
#plots the points for the above two variables with an alpha of 0.2 to allow for visualization of overlapping data points. A colour aesthetic was also used to show the diagnosis of each tumor.
  geom_point(aes(colour = diagnosis), alpha = 0.2) +
  
#sets the origin of the graph to 0,0 and ensures all points are included in the plot.
  expand_limits(x = 0, y = 0) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, max(cancer_sample$area_mean)+100)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(cancer_sample$perimeter_mean)+10)) +

#adds a linear trend line
  stat_smooth(method = 'lm', colour = "grey", se = FALSE, fullrange=TRUE) 
```

### Exercise 2.2

I decided to look at the P value so I used the tidy() function which provides the P value in column 5 of the generated tibble. I was second guessing myself because I've never seen a P value of 0, so to double check this I ran a Welch Two Sample t-test using the t.test() function. The results from t test are seen in the R console generated below and it shows that the p value is really small so tidy provided it as 0, meaning the results are very statistically significant.

```{r}
#provides data such as estimate, std.error, a statistic, and p.value
tidy(trendline)

#produces a Welch Two Sample t-test, provides statistical data including P value
x <- cancer_sample$area_mean
y <- cancer_sample$perimeter_mean
t.test(x, y)
```

# Exercise 3

### Exercise 3.1

##### Summary table from Milestone 2

```{r}
#selects cancer_sample dataset
(factored_mean_data <- cancer_sample %>%
  
#groups data based on the diagnosis variable
  group_by(diagnosis) %>%
  
#produces an additional variable with 3 possible categories as the observations based on area_mean
  mutate(relative_tumor_size = case_when(area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 TRUE ~ "large")) %>%
  
#only provides a tibble with the desired varibles
  select(ID,diagnosis,area_mean,relative_tumor_size,radius_mean:perimeter_mean,smoothness_mean:fractal_dimension_mean))

```

##### Produce a .csv file

```{r}
#writes the tibble 'factored_mean_data' intro a .csv file in the folder named 'output'
write_csv(factored_mean_data, here::here("~/Desktop/STAT_545/STAT_545A/mda-Kyle-Amaral/output", "fractored_mean_data.csv"))
```

### Exercise 3.2

The model object from Exercise 2 is named 'trendline'

```{r}

saveRDS(trendline, here::here("~/Desktop/STAT_545/STAT_545A/mda-Kyle-Amaral/output", "trendline.RDS"))


(readRDS("~/Desktop/STAT_545/STAT_545A/mda-Kyle-Amaral/output/trendline.RDS"))
```

# Additional Comments

My personal branch still doesn't work as it doesn't allow for pull requests to be made between the personal branch and main branch. I had to complete this whole Milestone on the main branch due to this issue. I continue to get an error and its not related to doing pull and pushes from Rstudio because both branches are up to date when I do so. 
