---
title: "mda-Milestone-3"
output: github_document
---

`````{r}
#loads needed packages to execute the code in this file and supresses startup messages
suppressPackageStartupMessages(library(datateachr)) #allows for access to the datasets in this file.
suppressPackageStartupMessages(library(tidyverse)) #loads packages needed for data visualization (ggplot2), and data manipulation (dplyr) that were used in this file.
suppressPackageStartupMessages(library(broom))
```

# Research Questions:

1. Is there a correlation between the area and perimeter of a tumor in regards to tumor diagnosis?
2. Is there a correlation between the tumor's fractal dimension and tumor diagnosis?

# Exercise 1

Had to make a new graph since none of my graphs in Milestone 2 had 3 factors. Below is the code to produce this original graph.

```{r}

cancer_sample %>%
  
  
  mutate(relative_tumor_size = case_when(area_mean < 200 ~ "very low",                                    
                                 area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ " very large")) %>%
  
  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
  
  geom_boxplot(width = 0.5) +
  
  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3)
```

#### Task Number: 1

```{r}

cancer_sample %>%
  
  
  mutate(relative_tumor_size = factor(case_when(area_mean < 200 ~ "very small",                                                                                        area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ "very large"),
                                 levels = c("very small", "small", "medium",  "large", "very large"))) %>%
  
  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
  
  geom_boxplot(width = 0.5) +
  
  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3)
```

I chose to do this ordering of the factors because I believe when looking at difference in sizes, it seems to be more easy on the eye when you go from the smallest value to the largest. It is also easier to view and see trends because its usually how this kind of measurements are ordered. For example, if I were to remove the axis labels and factor names, I would automatically think that in the original graph, the box plots went from smallest to largest, which it did not. 

#### Task Number: 2

```{r}

cancer_sample %>%
  
  
  mutate(relative_tumor_size = factor(case_when(area_mean < 200 ~ "very small",                                                                                        area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ "very large"),
                                 levels = c("very small", "small", "medium",  "large", "very large"))) %>%
  
  
  mutate(relative_tumor_size = fct_collapse(relative_tumor_size, 
              medium = "medium", large = "large", very_large = "very large", other = c("very small", "small"))) %>%

  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
  
  geom_boxplot(width = 0.5) + 
  
  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3) +
  
  
  scale_x_discrete(labels = c("other", "medium", "large", "very large"))
```

I chose to do this grouping because I was mostly interested in the medium to largest factors, but did want to remove any available data. By doing this group it also emphasizes how there are very little numbers of benign tumors, and they dont go past the medium factor. This provides information on tumor diagnosis and area_mean. 

# Exercise 2

### Exercise 2.0

Research Question: Is there a correlation between the area and perimeter of a tumor in regards to tumor diagnosis?
Variable of interest: area_mean

### Exercise 2.1
```{r}
trendline <- lm(perimeter_mean ~ area_mean, cancer_sample)
unclassed <- unclass(trendline)
print(unclassed)
```


```{r}
#selects cancer_sample dataset and produces a graph with area_mean on the x-axis and perimeter_mean on the y-axis.
cancer_sample %>%
  ggplot(aes(x = area_mean, y = perimeter_mean)) +
  
#plots the points for the above two variables with an alpha of 0.2 to allow for visualization of overlapping data points. A colour aesthetic was also used to show the diagnosis of each tumor.
  geom_point(aes(colour = diagnosis), alpha = 0.2) +
  
#sets the origin of the graph to 0,0 and ensures all points are included in the plot.
  expand_limits(x = 0, y = 0) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, max(cancer_sample$area_mean)+100)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(cancer_sample$perimeter_mean)+10)) +

#adds a linear trend line
  stat_smooth(method = 'lm', aes(colour = "red"), se = FALSE, fullrange=TRUE) +

#removes legend produced from trend line 
  theme(legend.position="none")
```

### Exercise 2.2

```{r}

tidy(trendline)

x <- cancer_sample$area_mean
y <- cancer_sample$perimeter_mean
t.test(x, y)
```

# Exercise 3

### Exercise 3.1

```{r}

```

### Exercise 3.2

```{r}

```

# Additional Comments
