---
title: "mda-Milestone-3"
output: github_document
---

`````{r}
#loads needed packages to execute the code in this file and supresses startup messages
suppressPackageStartupMessages(library(datateachr)) #allows for access to the datasets in this file.
suppressPackageStartupMessages(library(tidyverse)) #loads packages needed for data visualization (ggplot2), and data manipulation (dplyr) that were used in this file.
suppressPackageStartupMessages(library(broom))
```

# Research Questions:

1. Is there a correlation between the area and perimeter of a tumor in regards to tumor diagnosis?
2. Is there a correlation between the tumor's fractal dimension and tumor diagnosis?

# Exercise 1

Had to make a new graph since none of my graphs in Milestone 2 had 3 factors. Below is the code to produce this original graph.

```{r}

cancer_sample %>%
  
  
  mutate(relative_tumor_size = case_when(area_mean < 200 ~ "very low",                                    
                                 area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ " very large")) %>%
  
  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
  
  geom_boxplot(width = 0.5) +
  
  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3)
```

#### Task Number: 1

```{r}

cancer_sample %>%
  
  
  mutate(relative_tumor_size = factor(case_when(area_mean < 200 ~ "very small",                                                                                        area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ "very large"),
                                 levels = c("very small", "small", "medium",  "large", "very large"))) %>%
  
  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
  
  geom_boxplot(width = 0.5) +
  
  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3)
```

#### Task Number: 2

```{r}

cancer_sample %>%
  
  
  mutate(relative_tumor_size = factor(case_when(area_mean < 200 ~ "very small",                                                                                        area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ "very large"),
                                 levels = c("very small", "small", "medium",  "large", "very large"))) %>%
  
  
  mutate(relative_tumor_size = fct_collapse(relative_tumor_size, 
              medium = "medium", large = "large", very_large = "very large", other = c("very small", "small"))) %>%

  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
  
  geom_boxplot(width = 0.5) + 
  
  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3) +
  
  
  scale_x_discrete(labels = c("other", "medium", "large", "very large"))
```

# Exercise 2

### Exercise 2.0

Research Question: Is there a correlation between the area and perimeter of a tumor in regards to tumor diagnosis?
Variable of interest: area_mean

### Exercise 2.1
```{r}
trendline <- lm(perimeter_mean ~ area_mean, cancer_sample)
unclassed <- unclass(trendline)
print(unclassed)
```


```{r}
#selects cancer_sample dataset and produces a graph with area_mean on the x-axis and perimeter_mean on the y-axis.
cancer_sample %>%
  ggplot(aes(x = area_mean, y = perimeter_mean)) +
  
#plots the points for the above two variables with an alpha of 0.2 to allow for visualization of overlapping data points. A colour aesthetic was also used to show the diagnosis of each tumor.
  geom_point(aes(colour = diagnosis), alpha = 0.2) +
  
#sets the origin of the graph to 0,0 and ensures all points are included in the plot.
  expand_limits(x = 0, y = 0) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, max(cancer_sample$area_mean)+100)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(cancer_sample$perimeter_mean)+10)) +

#adds a linear trend line
  stat_smooth(method = 'lm', aes(colour = "red"), se = FALSE, fullrange=TRUE) +

#removes legend produced from trend line 
  theme(legend.position="none")
```

### Exercise 2.2

```{r}

tidy(trendline)

x <- cancer_sample$area_mean
y <- cancer_sample$perimeter_mean
t.test(x, y)
```

# Exercise 3

### Exercise 3.1

```{r}

```

### Exercise 3.2

```{r}

```

# Additional Comments
