---
title: "mda-Milestone-3"
output: github_document
---

`````{r}
#loads needed packages to execute the code in this file and supresses startup messages
suppressPackageStartupMessages(library(datateachr)) #allows for access to the datasets in this file.
suppressPackageStartupMessages(library(tidyverse)) #loads packages needed for data visualization (ggplot2), and data manipulation (dplyr) that were used in this file.
```

# Research Questions:

1. Is there a correlation between the area and perimeter of a tumor in regards to tumor diagnosis?
2. Is there a correlation between the tumor's fractal dimension and tumor diagnosis?

# Exercise 1

Had to make a new graph since none of my graphs in Milestone 2 had 3 factors. Below is the code to produce this original graph.

```{r}

cancer_sample %>%
  
  
  mutate(relative_tumor_size = case_when(area_mean < 200 ~ "very low",                                    
                                 area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ " very large")) %>%
  
  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
  
  geom_boxplot(width = 0.5) +
  
  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3)
```

#### Task Number: 1

```{r}

cancer_sample %>%
  
  
  mutate(relative_tumor_size = factor(case_when(area_mean < 200 ~ "very small",                                                                                        area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ "very large"),
                                 levels = c("very small", "small", "medium",  "large", "very large"))) %>%
  
  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
  
  geom_boxplot(width = 0.5) +
  
  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3)
```

#### Task Number: 2

```{r}

cancer_sample %>%
  
  
  mutate(relative_tumor_size = factor(case_when(area_mean < 200 ~ "very small",                                                                                        area_mean < 650 ~ "small",
                                 area_mean < 1000 ~ "medium",
                                 area_mean < 1500 ~ "large",
                                 TRUE ~ "very large"),
                                 levels = c("very small", "small", "medium",  "large", "very large"))) %>%
  
  
  mutate(relative_tumor_size = fct_collapse(relative_tumor_size, 
              medium = "medium", large = "large", very_large = "very large", other = c("very small", "small"))) %>%

  
  ggplot(aes(x = relative_tumor_size, y = area_mean)) +
  
  
  geom_boxplot(width = 0.5) + 
  
  
  geom_jitter(aes(colour = diagnosis), width = 0.2, size = 0.2, alpha = 0.3) +
  
  
  scale_x_discrete(labels = c("other", "medium", "large", "very large"))
```

# Exercise 2

### Exercise 2.0

Research Questions:
1.
2.

### Exercise 2.1

```{r}

```

### Exercise 2.2

```{r}

```

# Exercise 3

### Exercise 3.1

```{r}

```

### Exercise 3.2

```{r}

```

# Additional Comments
